---
title: "PeptideMHC2Prediction: Prediction of Peptide:MHC-II Molecules Interactions"
author: "Abdoelanser M. Degoot, Faraimunashe Chirove, and Wilfred Ndifon"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{PeptideMHC2Prediction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  
  output:
  knitr:::html_vignette:
    toc: yes
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## PeptideMHC2Prediction
An R package implementing biophysical methods for prediction of peptide and MHC-II molecules interactions, developed by Degoot and colleagues \url (https://www.frontiersin.org/articles/10.3389/fimmu.2018.01410/full). It predicts the probability of binding affinity  between  peptide of any length that greater than or equal 9 and 35 HLA-II molecules belonging to the three MHC-II loci of human genes: DRB, DP, and DQ. Use the __supportedMolecules( )__ function to see the molecules that covered by the current version of the package.

## Getting started
 First load the package
```{r}
##library(PeptideMHC2Prediction)

```

## Prediction for individual peptides

To predict the probability of binding affinity between peptide, for example say,
```{r}
pep <- "ADAGYAPATPAAAGA"
```
and an MHC-II molecule, for examples say DRB1_0401, based on trans-allelic  method, you can use the following: 
```{r}
##predictBinding(pep, "DRB1_0401", "DRB", 2)
```
Or using the intra-allelic method as follows:

```{r}
##predictBinding(pep, "DRB1_0401", "DRB", 1)
```
Please call this 

```{r}
# supportedMolecules()
```

function to see the 35 MHC-II molecules covered by the current version of the package.

The output of the function predictBinding is vector containg four items as follows:


```{r}
#   "ADAGYAPATPAAAGA", "DRB1_0401",x= "0.xxx", y= {0 or 1}
```

The first and second items are the peptide and the MHC-II molecule, respectively. The third item is a value for binding probability of the interaction between the peptide and MHC-II molecule as predicted by the model. The last item is a binarized value of the probability, it returns zero (0) if the probability less than 0.5 and 1 otherwise. 

##Prediction for set peptides

Instead of making predictions for individual peptides, the package also allows to make predictions for  set of peptides and MHC-II molecules, together organized in form of a dataframe of the three columns.  The columns of the dataframe must be in the following order: th peptide, the name of MHC-II moleclue, and the locus. For example, let us assume that we have a dataframe called __dataset__, which has the following structure. 

  Peptide | MHC | Locus |
|:---------------:|:--------:|:---------:|
|AAAGAEAGKATTEEQ| DRB1_0301|  DRB   |
|AFALVLLFCALASSC| DRB3_0101|  DRB   |
|AAAGAEAGKATTEEQ| DRB1_1501|  DRB   |
|SGLVWGQKYFKGNFQ| DPB10401|  DP   |
|SSNPTILSEGNSFTA| DQB10201|  DQ   |

Then you call the function __predictBindingSet__ to get prediction of the above dataset as follows:


```{r}
##predictBindingSet(dataset, opt = 1)
```
The parameter opt specifies the prediction method, 1 for intra-allelic method and 2 for trans-allelic method. The output of function __predictBindingSet__ is a dataframe of four columns, which has the below structure:

   Peptide | MHC | Probability |Binary Value|
|:---------------:|:--------:|:---------:|:---------:|
|AAAGAEAGKATTEEQ| DRB1_0301|  0.176   |  0  |
|AFALVLLFCALASSC| DRB3_0101|  0.549   |  1  |
|AAAGAEAGKATTEEQ| DRB1_1501|  0.679   |  1  |
|SGLVWGQKYFKGNFQ| DPB10401|   0.300  |  0 |
|SSNPTILSEGNSFTA| DQB10201|  0.026   |0|

## References
Abdoelnaser M. Degoot, Chirove Faraimunashe, and Ndifon Wilfred (2018)
Trans-Allelic Model for Prediction of Peptide:MHC-II Interactions, Frontiers in Immunology. URL:https://www.frontiersin.org/article/10.3389/fimmu.2018.01410.